---
title: "Initial Korean covid 19"
author: "Aylin Torres"
date: "11/12/24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

some of the data was uncoded for sex-sex must be just male or female 77 of the data enteries were coded sex as blank, so we omit these. state of a patient must be died or alive-we adjust that as well    


# METHODS

$h_0$ : There is no difference in dying depending on sex.

$h_0$ : There is a difference is dying depending on sex.

The Data is from korea in 2022.

result~sex 
this is catagorical ~ catagorical 
we will use helper file CAT_CAT



Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  



```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type= "percent")
```

  The bar chart shows more males died than women.

  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
 among the people who died there were 64% of men who died and 36% of women who died.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P-value of Chi-squared test is 0.001 and the fisheer exact test has p-value of .00009. The odd of the fisher test also taste that men are 2.3 more acceptable of dying from covid 19 therefore the p-value are lower than 0.05 we reject the null hypothesis.

## Conclusions

$H_0$: there is no difference in dying depending on sex.

$h_0$: there is difference in dying depending on sex.

The data is from korea in 2022.

result~sex
this is catagorical ~ catagorical
we will use helper fil CAT_CAT


the bar chart shows more males died than women.

Among the people who died there were 64% of men who died and 36% of women who died.

The P-value of Chi-squared test is 0.001 and the fisher exact test has a p-value of .00009. The odd of the fisher test also taste that men  are 2.3 more acceptable of dying from covid19 therefore the p-value are lower than 0.05 we reject the null hypothesis.






